---
title: "High_dim_flow_analysis"
output: html_document
date: '2022-10-03'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in packages

```{r}
library(BiocManager)
library(cytofWorkflow)
library(here)
library(tidyverse)
library(knitr)
```

# _____Cytek 25-color HNC PBMC pair_____

## • Loading & prepping data

```{r}
path <- '/Volumes/hqdinh2/Lab Members/Athena/Flow Experiments/Cytek_25_color_test_72537_73701_AEG_02072023/'

# saving path of flow file to data
my_flow_file <- paste(path, 'Unmixed/Multicolor/', sep = '')

# starting a random seed, saving markers as variable
set.seed(1333)

md <- read.csv(paste(path, 'Analysis/Dataframes/metadata.csv', sep = ''))

# need all four columns or this will NOT work
flow_data <- prepData(x = my_flow_file, md = md, transform = TRUE, cofactor = 150, FACS = TRUE)

saveRDS(flow_data, paste(path, 'Analysis/Objects/', Sys.Date(), '_flow_sce.RDS', sep = ''))
```

## • Clustering, setting maximum k value

```{r}
flow_data <- readRDS(paste(path, 'Analysis/Objects/', Sys.Date(), '_flow_sce.RDS', sep = ''))

flow_data <- flow_data[!rownames(flow_data) %like any% c('FSC%', 'SSC%', 'Time%', 'AF%')]

# ensuring we have removed any irrelevant markers
rownames(flow_data)

# below uses FlowSom clustering, k is a resolution term
flow_data <- CATALYST::cluster(flow_data, features = rownames(flow_data), xdim = 10, ydim = 10, maxK = 30, seed = 1333)

saveRDS(flow_data, paste(path, 'Analysis/Objects/', Sys.Date(), '_flow_sce.RDS', sep = ''))
```

## • Filter data based on expression

```{r}
# evaluate expression of markers for filtering
plot(density(flow_data@assays@data$exprs['CD45',]))
plot(density(flow_data@assays@data$exprs['ViaDye Red-A',]))

flow_data <- flow_data[, flow_data@assays@data$exprs['CD45',] > 0]
flow_data <- flow_data[, flow_data@assays@data$exprs['ViaDye Red-A',] < 2]
```

## • UMAP

### •• Run UMAP
```{r}
# subsetting so we do not include unnecessary markers
flow_data <- flow_data[!rownames(flow_data) %like any% c('FSC%', 'SSC%', 'Time%', 'AF%', 'ViaDye%')]
flow_data <- flow_data[!rownames(flow_data) == 'CD45']

flow_data <- runDR(flow_data, "UMAP", features = rownames(flow_data), assay = "exprs")

delta_area(flow_data)
```

### •• Plot UMAP
```{r}
plotDR(flow_data, "UMAP", color_by = rownames(flow_data), scale = TRUE, ncol = 5)

png(paste(path, '/Analysis/Figures/UMAP_splitbycondition.png', sep = ''), units = 'in', width = 5, height = 4, res = 300)
plotDR(flow_data, 'UMAP', color_by = 'meta6', facet = 'condition')
dev.off()
```


## • Cluster heatmap

```{r}
library(data.table)
library(dplyr)
library(DescTools)

png(filename = paste(path, 'Analysis/Figures/cluster_marker_heatmap.png', sep = ''), units = 'in', width = 6, height = 4, res = 300)
plotExprHeatmap(flow_data, k = "meta6", by = "cluster_id")
dev.off()
```


## • Frequency boxplots

```{r}
png(paste(path, '/Analysis/Figures/panel1_boxplots_splitbycluster.png', sep = ''), units = 'in', width = 7, height = 5, res = 300)
plotAbundances(flow_data, k = "meta6", by = "cluster_id") + scale_fill_manual(values=c('Green', 'Black', 'Blue', 'Red')) + scale_color_manual(values = c('Blue', 'Red'))
dev.off()
```

## • Diagnostic plots

### •• Marker density plots
```{r}
p <- plotExprs(flow_data, color_by = "condition")
p$facet$params$ncol <- 6
png(paste(path, '/Analysis/Figures/DIAGNOSTIC_marker_densities.png', sep = ''), units = 'in', width = 7, height = 5, res = 300)
p
dev.off()
```

### •• Assessing # of cells
```{r}
n_cells(flow_data)
# 72537 73701 
# 3199  1077 
```

### •• Cell counts per sample
```{r}
colData(flow_data)$sample_id <- as.factor(colData(flow_data)$sample_id)
colData(flow_data)$sample_id <- as.character(colData(flow_data)$sample_id)

png(paste(path, '/Analysis/Figures/DIAGNOSTIC_sample_counts.png', sep = ''), units = 'in', width = 3.5, height = 5, res = 300)
plotCounts(flow_data, group_by = "sample_id", color_by = "condition")
dev.off()
```

### •• MDS plot
```{r}
plotExprHeatmap(flow_data, scale = "last", hm_pal = rev(hcl.colors(10, "YlGnBu")))
```


# _____Mouse neutrophil PANEL1_____

## • Loading data

```{r}
# saving path of flow file to data
my_flow_file <- '/Volumes/oncology/Athena_Dinh/exported CD45+ single live cells panel 1/'

# starting a random seed, saving markers as variable
set.seed(1333)
```

## • Prepping data

```{r}
md <- read.csv('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Mouse_ICI_TRT_flow_metadata_panel1.csv')

flow_data <- prepData(x = my_flow_file, md = md, transform = TRUE, cofactor = 150, FACS = TRUE)

saveRDS(flow_data, '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/flow_clustering_singlecellexperiment_panel1.RDS')
```

## • Clustering, setting maximum k value

```{r}
flow_data <- readRDS('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/flow_clustering_singlecellexperiment_panel1.RDS')

flow_data <- flow_data[!rownames(flow_data) %like any% c('FSC%', 'SSC%', 'LD%', 'Time%', 'CD45%')]

rownames(flow_data)[rownames(flow_data) %like% 'CD8%'] <- 'CD8'
rownames(flow_data)[rownames(flow_data) %like% 'CD45%'] <- 'CD45'
rownames(flow_data)[rownames(flow_data) %like% 'F4_80%'] <- 'F4_80'
rownames(flow_data)[rownames(flow_data) %like% 'CD11b%'] <- 'CD11b'
rownames(flow_data)[rownames(flow_data) %like% 'CD19%'] <- 'CD19'
rownames(flow_data)[rownames(flow_data) %like% 'CD4%'] <- 'CD4'
rownames(flow_data)[rownames(flow_data) %like% 'GR1%'] <- 'GR1'
rownames(flow_data)[rownames(flow_data) %like% 'CD11C%'] <- 'CD11c'

rownames(flow_data)

# below uses FlowSom clustering, k is a resolution term
flow_data <- CATALYST::cluster(flow_data, features = rownames(flow_data), xdim = 10, ydim = 10, maxK = 30, seed = 1333)

saveRDS(flow_data, '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/flow_clustering_singlecellexperiment_panel1.RDS')
```

## • Consensus Cluster Decision

```{r}
delta_area(flow_data)
```

## • Cluster heatmap

```{r}
library(data.table)
library(dplyr)
library(DescTools)

flow_mini <- scrabbitr::downsampleSCE(sce = flow_data, ncells = 20000)

saveRDS(flow_mini, file = '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/MINI_flow_clustering_singlecellexperiment_panel1.RDS')

png(filename = '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/cluster_marker_heatmap_panel1.png', units = 'in', width = 6, height = 4, res = 300)
plotExprHeatmap(flow_mini, k = "meta8", by = "cluster_id")
dev.off()
```

## • UMAP

### •• Run UMAP

```{r}
flow_mini <- runDR(flow_mini, "UMAP", features = rownames(flow_mini), assay = "exprs")
```

### •• Plot UMAP (all markers)

```{r}
plotDR(flow_mini, "UMAP", color_by = rownames(flow_mini), scale = TRUE, ncol = 5)

png('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/AEG_Ultracomp_gating_Panel1/panel1_UMAP_splitbycondition.png', units = 'in', width = 5, height = 4, res = 300)
plotDR(flow_mini, 'UMAP', color_by = 'meta8', facet = 'condition', k_pal = color_clusters)
dev.off()
```

## • Frequency boxplots

```{r}
png('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/AEG_Ultracomp_gating_Panel1/panel1_boxplots_splitbycluster.png', units = 'in', width = 7, height = 5, res = 300)
plotAbundances(flow_mini, k = "meta8", by = "cluster_id") + scale_fill_manual(values=c('Green', 'Black', 'Blue', 'Red')) + scale_color_manual(values = c('Green', 'Black', 'Blue', 'Red'))
dev.off()
```

# _____Mouse neutrophil PANEL 2_____

# Loading and prepping data

```{r}
# saving path of flow file to data
my_flow_file <- '/Volumes/oncology/Athena_Dinh/exported CD45+ single live cells panel 2/'

# starting a random seed, saving markers as variable
set.seed(1333)
```

# Prepping data

```{r}
md <- read.csv('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Mouse_ICI_TRT_flow_metadata_panel2.csv')

flow_data <- prepData(x = my_flow_file, md = md, transform = TRUE, cofactor = 150, FACS = TRUE)
```

# Clustering, setting maximum k value

```{r}
flow_data <- readRDS('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/flow_clustering_singlecellexperiment_panel2.RDS')

set.seed(1333)

# below uses FlowSom clustering, k is a resolution term
flow_data <- CATALYST::cluster(flow_data, features = rownames(flow_data), xdim = 10, ydim = 10, maxK = 30, seed = 1333)

saveRDS(flow_data, '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/flow_clustering_singlecellexperiment_panel2.RDS')
```

# Consensus Cluster Decision

```{r}
delta_area(flow_data)
```

# Cluster heatmap

```{r}
library(data.table)
library(dplyr)
library(DescTools)

flow_mini <- scrabbitr::downsampleSCE(sce = flow_data, ncells = 20000)

flow_mini <- readRDS(file = '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/MINI_flow_clustering_singlecellexperiment_panel2.RDS')

saveRDS(flow_mini, file = '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/MINI_flow_clustering_singlecellexperiment_panel2.RDS')

png(filename = '/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/cluster_marker_heatmap.png', units = 'in', width = 6, height = 4, res = 300)
# change meta column to change resolution (meta5, meta10, meta15, ...)
plotExprHeatmap(flow_mini, k = "meta12", by = "cluster_id")
dev.off()
```

# UMAP

## Run UMAP

```{r}
flow_mini <- runDR(flow_mini, "UMAP", features = rownames(flow_mini), assay = "exprs")
```

## Plot UMAP (all markers)

```{r}
plotDR(flow_mini, "UMAP", color_by = rownames(flow_mini), scale = TRUE, ncol = 5)

png('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/AEG_Ultracomp_gating_Panel2/panel2_UMAP_splitbycondition.png', units = 'in', width = 5, height = 4, res = 300)
plotDR(flow_mini, 'UMAP', color_by = 'meta12', facet = 'condition', k_pal = color_clusters)
dev.off()
```

## Frequency boxplots

```{r}
png('/Volumes/hqdinh2/Projects/HNC_SPORE/Mouse_Neutrophil_ICITRT/Round1/AEG_Ultracomp_gating_Panel2/panel2_boxplots_splitbycluster.png', units = 'in', width = 7, height = 5, res = 300)
plotAbundances(flow_mini, k = "meta12", by = "cluster_id") + scale_fill_manual(values=c('Green', 'Black', 'Blue', 'Red')) + scale_color_manual(values = c('Green', 'Black', 'Blue', 'Red'))
dev.off()
```

## 
